---
title: "Reverse Titanic"
output: html_document
---


```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(randomForest)
library(mice)
library(e1071)
library(gbm)
```

Importing data, making sure to import blank character data as "NA"s rather than just "" since R thinks these are characters but really they are missing data
```{r}
train <- read.csv("Titanic/Data/titanic_train.csv", na.strings = c("", "NA"), stringsAsFactors = F)
test <- read.csv("Titanic/Data/titanic_test.csv", na.strings = c("", "NA"), stringsAsFactors = F)
full <- bind_rows(train, test)
```

#Missing values
First we see how many values are missing and from which columns
```{r}
head(train)
```

```{r}
for(i in 1:length(names(train))) {
  print(sum(is.na(train[,i])))
}
```
Train is missing 177 age values, 687 cabin values, and 2 embarked values. 
```{r}
for(i in 1:length(names(test))) {
  print(sum(is.na(test[,i])))
}
```
Test is missing 86 age values, 1 fare value, and 327 cabin values. 

Since the size of our data is somewhat small, we cannot just remove entire rows which are missing data. So we have to put in some sensible replacements.

First we start with the 2 missing embarked values in train.
```{r}
missing_embarked_ind <- which(is.na(train$Embarked))
train[missing_embarked_ind,]
```
The most likely columns which correlate to which port they embarked are fare and passenger class. The fares were 80 and 80 and the Pclass's were 1 and 1. 

We are going to visulize fare, embarked, and p class in one graph.
```{r}
missing_embar <- train %>%
  filter(PassengerId != 62, PassengerId != 830)

ggplot(missing_embar, aes(x = Embarked, y = Fare, fill = factor(Pclass))) +
  geom_boxplot() +
  geom_hline(aes(yintercept=80), 
    colour='red', linetype='dashed', lwd=2)
```
From the looks of things, a passenger with a Fare of $80 and a P class of 1 was most likely from C. Thus we'll fill in those NA's with C. 

```{r}
train$Embarked[missing_embarked_ind] <- "C"
full$Embarked[missing_embarked_ind] <- "C"
```

Now for the missing fare value in the test data.
```{r}
missing_fare_ind <- which(is.na(test$Fare))
test[missing_fare_ind,]
```
This passenger was in class 3 and embarked from S. Similar to before, we are going to visualize all passengers who were class 3 and embarked from S and see what a good value to stick in is.

```{r}
ggplot(full[full$Pclass == '3' & full$Embarked == 'S',],
  aes(x = Fare)) +
  geom_density() +
  geom_vline(aes(xintercept = median(Fare, na.rm = T)),
                 color = 'red', linetype = 'dashed', lwd = 1)

```
Thus it seems ok to replace the missing value with the median value.
```{r}
temp_median_fare <- median(full[full$Pclass == '3' & full$Embarked == 'S',]$Fare , na.rm = T)
test[missing_fare_ind,]$Fare <- temp_median_fare
full[missing_fare_ind,]$Fare <- temp_median_fare
```


Now let's start filling in those missing age values. We are going to use the mice package which uses some markov chain based algorithim to fill in the data...or something. 
```{r}
#make certain variables factors
factor_vars <- c('PassengerId','Pclass','Sex','Embarked')
full[factor_vars] <- lapply(full[factor_vars], factor)
```

```{r include = FALSE, message = FALSE, warnings = FALSE}
#perform mice imputation excluding less useful variables 
mice_mod <- mice(full[, !names(full) %in% c('PassengerId','Name','Ticket','Cabin')], method='rf')
#save the complete output
mice_output <- complete(mice_mod)
```
Now let's check the results to see if the made up ages seem okay.
```{r}
par(mfrow=c(1,2))
hist(full$Age, freq=F, main='Age: Original Data', 
  col='darkgreen', ylim=c(0,0.04))
hist(mice_output$Age, freq=F, main='Age: MICE Output', 
  col='lightgreen', ylim=c(0,0.04))
```
They do so we can replace the age vector with the mice model output.

```{r}
full$Age <- mice_output$Age
```

Now we split the full set back into training and test.
```{r}
train <- full[1:891,]
test <- full[892:1309,]
```

#Random Forest
Build model using random forest
```{r}
rf_model <- randomForest(factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, data = train)

#show model error
plot(rf_model, ylim=c(0,0.36))
legend('topright', colnames(rf_model$err.rate), col=1:3, fill=1:3)

```
Black line is overall error rate, red is for died, and green is for survived. 

Now let's check variable importance.
```{r}
importance <- importance(rf_model)
varImportance <- data.frame(Variables = row.names(importance), 
                            Importance = round(importance[ ,'MeanDecreaseGini'],2))

# Create a rank variable based on importance
rankImportance <- varImportance %>%
  mutate(Rank = paste0('#',dense_rank(desc(Importance))))

# Use ggplot2 to visualize the relative importance of variables
ggplot(rankImportance, aes(x = reorder(Variables, Importance), 
    y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
    hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip()
```

Now we make our prediction with random forest.
```{r}
prediction_rf <- predict(rf_model,test)
```


#SVM 
Use SVM (good with small datasets)
```{r}
svm_model <- train(factor(Survived) ~ Sex + Fare + Age + Pclass + SibSp + Parch + Embarked, data = train, method = 'svmRadial')
prediction_SVM <- predict(svm_model, test)
```


#GBM
Got to use 1 boosting algorithim somewhere
```{r}
boost_model <- train(factor(Survived) ~ Sex + Fare + Age + Pclass + SibSp + Parch + Embarked, data = train, method = 'gbm')
prediction_GBM <- predict(boost_model, test)
```

Ensemble predictions:
```{r}
ensemble <- data.frame(PassengerID = test$PassengerId, RandomForest = as.vector(prediction_rf), SVM = as.vector(prediction_SVM), GBM = as.vector(prediction_GBM))

survived <- c()
for(i in 1:nrow(ensemble)){
  if(as.numeric(ensemble$RandomForest[i])-1 + as.numeric(ensemble$SVM[i])-1 + as.numeric(ensemble$GBM[i])-1 <= 1) {
    survived[i] = 0
  } else {
    survived[i] = 1
  }
}
```

```{r}
solution <- data.frame(PassengerID = test$PassengerId, Survived = survived)

#write to csv file
write.csv(solution, file = 'rf_titanic_solution', row.names = F)

```
















